Preventing Runtime Errors of Redis at Compile Time

### 2015-11-20 paper.tex
errata

\paragraph{The Cause} Every key is associated with a value, and every value has
 it's own type. But Most commands in Redis only work with a certain type of
 value. When a command is used on a wrong type of key, a runtime error occurs.
 The problems illustrated above arise from the absence of type checking, with
 respects to \textbf{the type of a value that associates with a key}.
 These problems could have been avoided, if we could know the type every key
 associates with in advance, and prevent programs with invalid commands from
 executing.

@@

We began with making Hedis a dynamically checked embedded DSL, and implemented a
 runtime type checker that keeps track of types of each variable. But then we
 found that things can be a lot easier, by leveraging the host language's type
 checker. We encode variable bindings with \emph{type-level lists} and
 \emph{strings}, and control the effects on the bindings with
 \emph{indexed monad}. In contrast to the former approach, we \textbf{embedded}
 our type checker into Haskell's type system, without having to build a
 \textbf{standalone} one on the term level.

@@

Luckily, with recently added GHC extension \emph{data kinds}, suitable
 datatype will be automatically promoted to be a kind, and its value
 constructors to be type constructors.

@@

Give rise to the following kinds and type constructors:\footnotemark\footnotemark

\footnotetext{To distinguish between types and promoted constructors that have
 ambiguous names, prefix promoted constructor with a single quote like
 \mintinline{text}{'Nil} and \mintinline{text}{'Cons}}
\footnotetext{All kinds have \emph{sort} BOX in Haskell\cite{sorts}}

@@

Instead of letting users writing these boilerplates, we will be doing these
 serializations/deserializations for them. With the help from
 \mintinline{text}{cereal}, a binary serialization library.
 \mintinline{text}{cereal} comes with these two functions:

@@

In other words, we require that the key to be associated with a list,
 textbf{unless} it doesn't exist at all.

@@

With \emph{constraint kind}, a recent addition to GHC. Type constraints now has
 its own kind: \mintinline{haskell}{Constraint}. That means type constraints
 are not restricted to the left side of a \mintinline{haskell}{=>} anymore,
 they could appear in anywhere that accepts something of kind
 \mintinline{haskell}{Constraint}, and any type that has kind
 \mintinline{haskell}{Constraint} can also be used as a type constraint.
 \footnote{See \url{https://downloads.haskell.org/~ghc/7.4.1/docs/html/users_guide/constraint-kind.html}.}

@@

At the point of writing, transactions are not supported in our implementation.
 We are planning to add it in the future, and we are expecting that there
 wouldn't be much difficulties, since we've implemented a runtime type checker
 specifically targeting Redis transactions once, before we moved on to the
 types.

@@

While Redis can be viewed as a non-relational database system (although Redis
 seems reluctant to admit this in recent years), there also similar goals on
 relational database systems, trying to achieve a safer database interface by
 making them statically checked.
 \emph{HaskellDB}\cite{haskelldb, haskelldbimproved} expresses quires
 with relational algebra-like combinators, wrapped in phantom types. Or as
 examples, at first to demonstrate the power of dependent types in
 Agda\cite{pi}, and then with singletons in Haskell\cite{singletons}.

@@

\paragraph{Acknowledgements}
This material is based upon work supported by Institute of Information Science,
Academia Sinica under the Summer Internship Program of 2015.

>>>>

\paragraph{The Cause} Every key is associated with a value, and every value has
 it's own type. But most commands in Redis only work with a certain type of
 value. When a command is used on a wrong type of key, a runtime error occurs.
 The problems illustrated above arise from the absence of type checking, with
 respects to \textbf{the type of a value that associates with a key}.
 These problems could have been avoided, if we could know the type every key
 associates with in advance, and prevent programs with invalid commands from
 executing.

@@

We began with making Hedis a dynamically typechecked embedded DSL, and implemented a
 runtime type checker that keeps track of types of all the variable. But then we
 found that things can be a lot easier, by leveraging the host language's type
 checker. We encode variable bindings with \emph{type-level lists} and
 \emph{strings}, and control the effects on the bindings with
 \emph{indexed monad}. In contrast to the former approach, we \textbf{embedded}
 our type checker into Haskell's type system, without having to build a
 \textbf{standalone} one on the term level.

@@

Luckily, with recently added GHC extension \emph{data kinds}, suitable
 datatype will be automatically promoted to be a kind, and its value
 constructors to be type constructors. The following type {\mintinline{haskell}{List}

@@

Give rise to the following kinds and type constructors:\footnote{To distinguish
 between types and promoted constructors that have
 ambiguous names, prefix promoted constructor with a single quote like
 \mintinline{text}{'Nil} and \mintinline{text}{'Cons}}
\footnote{All kinds have \emph{sort} BOX in Haskell\cite{sorts}}

@@

Instead of letting users writing these boilerplates, we can do these
 serializations/deserializations for them. With the help from
 \mintinline{text}{cereal}, a binary serialization library.
 \mintinline{text}{cereal} comes with these two functions:

@@

In other words, we require that the key to be associated with a list,
 \textbf{unless} it doesn't exist at all.

@@

With \emph{constraint kind}, a recent addition to GHC, type constraints now has
 its own kind: \mintinline{haskell}{Constraint}. That means type constraints
 are not restricted to the left side of a \mintinline{haskell}{=>} anymore,
 they could appear in anywhere that accepts something of kind
 \mintinline{haskell}{Constraint}, and any type that has kind
 \mintinline{haskell}{Constraint} can also be used as a type constraint.
 \footnote{See \url{https://downloads.haskell.org/~ghc/7.4.1/docs/html/users_guide/constraint-kind.html}.}

@@

At this point of writing, transactions are not supported in our implementation.
 We are planning to add it in the future, and we are expecting that there
 wouldn't be much difficulty, since we've implemented a runtime type checker
 specifically targeting Redis transactions once, before we moved on to the
 types.

@@

While Redis can be viewed as a non-relational database system (although Redis
 seems reluctant to admit this in recent years), there also are similar goals on
 relational database systems, trying to achieve a safer database interface by
 making them statically checked.
 \emph{HaskellDB}\cite{haskelldb, haskelldbimproved} expresses quires
 with relational algebra-like combinators, wrapped in phantom types. Or as
 examples, at first to demonstrate the power of dependent types in
 Agda\cite{pi}, and then with singletons in Haskell\cite{singletons}.

@@

\paragraph{Acknowledgements}
\footnote{Hidden for the purposes of double-blind reviewing}
